### 2022-09-29 Final_Project_Work_HW2.Rmd
Done through residuals code for final fit

---
title: "Final_Project_Work_HW2_JaJuArEm"
author: "Jacob Kresnicka"
date: "2022-09-21"
output: html_document
---

# library statements

```{r}
library(ISLR)
library(dplyr)
library(readr)
library(broom)
library(ggplot2)
library(tidymodels) 
library(stringr)
library(splitstackshape)
tidymodels_prefer()
```

# read in data

```{r}
library(readr)
imdb_top_1000 <- read_csv("~/Desktop/Statistical Machine Learning/R Files/Final Project/imdb_top_1000.csv")
head(imdb_top_1000)
```

# data cleaning

```{r}
imdb_clean <- imdb_top_1000 %>%
  select(-Poster_Link, -Certificate) %>%
  cSplit("Genre", sep = ",", direction = "wide")

runtime_clean <- imdb_top_1000$Runtime %>%
  str_replace(" min", "") %>%
  as.numeric()

imdb_clean$Runtime <- runtime_clean
```

# creation of cv folds

```{r}
data_cv10 <- vfold_cv(imdb_clean, v = 10)
```

# model spec

```{r}
lm_spec <-
    linear_reg() %>% 
    set_engine(engine = 'lm') %>% 
    set_mode('regression')
```

# recipes & workflows

```{r}
full_rec <- recipe(Gross ~ ., data = imdb_clean) %>%
    step_rm(Series_Title, Released_Year, Overview,
            Director, Star1, Star2, Star3, Star4, Genre_2, Genre_3) %>% #variables causing issues
    step_nzv(all_predictors()) %>% 
    step_normalize(all_numeric_predictors()) %>% 
    step_dummy(all_nominal_predictors()) %>%
    step_naomit(Gross)

full_lm_wf <- workflow() %>%
    add_recipe(full_rec) %>%
    add_model(lm_spec)

full_model <- fit(full_lm_wf, data = imdb_clean)

full_model %>% tidy()
```

# fit & tune models

# LASSO

```{r}

# Lasso Model Spec with tune
lm_lasso_spec_tune <- 
  linear_reg() %>%
  set_args(mixture = 1, penalty = tune()) %>% ## mixture = 1 indicates Lasso
  set_engine(engine = 'glmnet') %>% #note we are using a different engine
  set_mode('regression')

# Recipe with standardization (!) --> just include all these always
data_rec <- recipe(Gross ~ ., data = imdb_clean) %>%
    step_rm(Series_Title, Released_Year, Overview,
            Director, Star1, Star2, Star3, Star4, Genre_2, Genre_3) %>%
    step_nzv(all_predictors()) %>% # removes variables with the same value (so duplicates don't mess up model)
    step_novel(all_nominal_predictors()) %>% # important if you have rare categorical variables 
    step_normalize(all_numeric_predictors()) %>%  # super important standardization step for LASSO
    step_dummy(all_nominal_predictors()) %>%  # creates indicator variables for categorical variables
    step_naomit(Gross)

# Workflow (Recipe + Model)
lasso_wf_tune <- workflow() %>% 
  add_recipe(data_rec) %>%
  add_model(lm_lasso_spec_tune)

# Tune Model (trying a variety of values of Lambda penalty)
penalty_grid <- grid_regular(
  penalty(range = c(-1, 8)),
  levels = 60)

tune_res <- tune_grid( # new function for tuning parameters
  lasso_wf_tune, # workflow
  resamples = data_cv10, # cv folds
  metrics = metric_set(rmse, mae),
  grid = penalty_grid # penalty grid defined above
)

# Visualize Model Evaluation Metrics from Tuning
autoplot(tune_res) + theme_classic()

# Summarize Model Evaluation Metrics (CV)
collect_metrics(tune_res) %>%
  filter(.metric == 'rmse') %>% # or choose mae
  select(penalty, rmse = mean)

best_penalty <- select_best(tune_res, metric = 'rmse') # choose penalty value based on lowest mae or rmse

# Fit Final Model
final_wf <- finalize_workflow(lasso_wf_tune, best_penalty) # incorporates penalty value to workflow

final_fit <- fit(final_wf, data = imdb_clean)

tidy(final_fit)
```

```{r}
best_penalty <- select_best(tune_res, metric = 'mae') # choose penalty value based on lowest cv mae
best_penalty

# Use this if worried about overfitting
best_se_penalty <- select_by_one_std_err(tune_res, metric = 'mae', desc(penalty)) # choose largest penalty value within 1 se of the lowest cv mae
best_se_penalty
```

```{r}
final_wf <- finalize_workflow(lasso_wf_tune, best_penalty) # incorporates penalty value to workflow
final_wf_se <- finalize_workflow(lasso_wf_tune, best_se_penalty) # incorporates penalty value to workflow

final_fit <- fit(final_wf, data = imdb_clean)
final_fit_se <- fit(final_wf_se, data = imdb_clean)

tidy(final_fit)
tidy(final_fit_se)
```

```{r}
final_fit_se %>% tidy() %>% filter(estimate != 0)
```

```{r}
lasso_mod_out <- final_fit_se %>%
    predict(new_data = imdb_clean) %>%
    bind_cols(imdb_clean) %>%
    mutate(resid = Gross - .pred)

ggplot(lasso_mod_out, aes(x = .pred, y = resid)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") + 
    theme_classic()

ggplot(lasso_mod_out, aes(x = Runtime, y = resid)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") + 
    theme_classic()

ggplot(lasso_mod_out, aes(x = IMDB_Rating, y = resid)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") + 
    theme_classic()

ggplot(lasso_mod_out, aes(x = No_of_Votes, y = resid)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") + 
    theme_classic()
```

2.

3. Analyzing and predicting gross revenue of movies could have harmful effects on the performance of movies in box offices. For example, if we were to use our model to predict the success of a new movie, our model's prediction would affect people's decision making in whether to see the movie. This could affect the actual outcome of how movies perform at the box office. Our model's predictions could introduce bias into people's decisions of whether or not to see a movie.

The main caution we want to keep in mind when communicating our work is that our model may be bias against new names in the movie industry. The model cannot determine a director, actor, or actresses' talent. It only bases its predictions off of how previous movies that individual was involved in previously performed.