### 2022-10-16 Final_Project_Work_HW2.Rmd
HW3 edits

---
title: 'STAT 253 Final Project'
author: "Jacob, Julian, Aristo, and Emily"
date: "September 30, 2022"
output: 
  html_document:
    df_print: paged
    code_download: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE, message=FALSE, warning=FALSE)
```

# Library Statements

```{r}
library(ISLR)
library(dplyr)
library(readr)
library(broom)
library(ggplot2)
library(tidymodels) 
library(stringr)
library(splitstackshape)
library(tidymodels)
library(lubridate)
library(rpart.plot)
library(cluster)
library(forcats)
tidymodels_prefer()
```

# Read in Data

```{r}
library(readr)
imdb_top_1000 <- read_csv("~/Desktop/Statistical Machine Learning/R Files/Final Project/imdb_top_1000.csv")
head(imdb_top_1000)
```

# Data Cleaning

```{r}
imdb_clean <- imdb_top_1000 %>%
  select(-Poster_Link, -Certificate) %>%
  cSplit("Genre", sep = ",", direction = "wide") %>%
  mutate(Gross = log(Gross))

runtime_clean <- imdb_top_1000$Runtime %>%
  str_replace(" min", "") %>%
  as.numeric()

imdb_clean$Runtime <- runtime_clean
```

# Creation of CV Folds

```{r}
data_cv10 <- vfold_cv(imdb_clean, v = 10)
```

# Model Spec for Linear Regression Model

```{r}
lm_spec <-
    linear_reg() %>% 
    set_engine(engine = 'lm') %>% 
    set_mode('regression')
```

# Recipes and Workflows

```{r}
full_rec <- recipe(Gross ~ ., data = imdb_clean) %>%
    step_rm(Series_Title, Released_Year, Overview,
            Director, Star1, Star2, Star3, Star4, Genre_2, Genre_3) %>% #variables causing issues
    step_nzv(all_predictors()) %>% 
    step_normalize(all_numeric_predictors()) %>% 
    step_dummy(all_nominal_predictors()) %>%
    step_naomit(Gross)

full_lm_wf <- workflow() %>%
    add_recipe(full_rec) %>%
    add_model(lm_spec)

full_model <- fit(full_lm_wf, data = imdb_clean)

full_modelcv <- fit_resamples(full_lm_wf, resamples = data_cv10, metrics = metric_set(rmse, rsq, mae))

full_model %>% tidy()
```

# LASSO - Fit and Tune Models

```{r}

# Lasso Model Spec with tune
lm_lasso_spec_tune <- 
  linear_reg() %>%
  set_args(mixture = 1, penalty = tune()) %>% ## mixture = 1 indicates Lasso
  set_engine(engine = 'glmnet') %>% #note we are using a different engine
  set_mode('regression')

# Recipe with standardization (!) --> just include all these always
data_rec <- recipe(Gross ~ ., data = imdb_clean) %>%
    step_rm(Series_Title, Released_Year, Overview,
            Director, Star1, Star2, Star3, Star4, Genre_2, Genre_3) %>%
    step_nzv(all_predictors()) %>% # removes variables with the same value (so duplicates don't mess up model)
    step_novel(all_nominal_predictors()) %>% # important if you have rare categorical variables 
    step_normalize(all_numeric_predictors()) %>%  # super important standardization step for LASSO
    step_dummy(all_nominal_predictors()) %>%  # creates indicator variables for categorical variables
    step_naomit(Gross)

# Workflow (Recipe + Model)
lasso_wf_tune <- workflow() %>% 
  add_recipe(data_rec) %>%
  add_model(lm_lasso_spec_tune)

# Tune Model (trying a variety of values of Lambda penalty)
penalty_grid <- grid_regular(
  penalty(range = c(-3, 1)),
  levels = 30)

tune_res <- tune_grid( # new function for tuning parameters
  lasso_wf_tune, # workflow
  resamples = data_cv10, # cv folds
  metrics = metric_set(rmse, mae),
  grid = penalty_grid # penalty grid defined above
)

# Visualize Model Evaluation Metrics from Tuning
autoplot(tune_res) + theme_classic()
```

# Collect CV Metrics and Select Best Model

```{r}

# Summarize Model Evaluation Metrics (CV)
lasso_mod <- collect_metrics(tune_res) %>%
  filter(.metric == 'rmse') %>% # or choose mae
  select(penalty, rmse = mean)

best_penalty <- select_best(tune_res, metric = 'rmse') # choose penalty value based on lowest mae or rmse

lasso_mod
```

# Fit Final Model

```{r}

# Fit Final Model
final_wf <- finalize_workflow(lasso_wf_tune, best_penalty) # incorporates penalty value to workflow

final_fit <- fit(final_wf, data = imdb_clean)

lasso_fit <- fit_resamples(final_wf, resamples = data_cv10, metrics = metric_set(rmse, rsq, mae))

tidy(final_fit)
```

```{r}
# Final ("best") model predictors and coefficients

final_fit %>% tidy() %>% filter(estimate != 0)
```

# Visualize Residuals

```{r}

# Visualize Residuals
lasso_mod_out <- final_fit %>%
    predict(new_data = imdb_clean) %>%
    bind_cols(imdb_clean) %>%
    mutate(resid = Gross - .pred)

ggplot(lasso_mod_out, aes(x = .pred, y = resid)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") + 
    theme_classic()

ggplot(lasso_mod_out, aes(x = Runtime, y = resid)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") + 
    theme_classic()

ggplot(lasso_mod_out, aes(x = IMDB_Rating, y = resid)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") + 
    theme_classic()

ggplot(lasso_mod_out, aes(x = No_of_Votes, y = resid)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") + 
    theme_classic()
```

The overall goal of our linear regression model is to predict a movie's gross box office revenue based on a number of predictors. The goal of our initial investigation was to determine which subset of predictors are the most important for inclusion in our model. In order to determine the "best" linear regression model to achieve our analysis goals, we utilized the LASSO algorithm. The goal of LASSO is to decrease variance in our model by constraining the coefficient estimates and discouraging the inclusion of weakly informative predictors through the use of a tuning parameter. This improves the predictive accuracy of our model. We chose this algorithm over the best subset selection algorithm because LASSO is substantially more computationally efficient. Furthermore, since our main analysis goal is predictive accuracy, we decided against using forward stepwise or backwards stepwise selection because these algorithms do not guarantee the best overall model. A model using a smaller subset of variables may have increased interpretability, but this would come at the expense of predictive accuracy.

The main caution we want to keep in mind when communicating our work is that our model may be bias against new names in the movie industry. The model cannot determine a director, actor, or actresses' talent. It only bases its predictions off of how previous movies that individual was involved in previously performed. Furthermore, the inclusion of the number of votes predictor in our model raises further concerns of potential bias. Our initial evaluation of the residual vs. predictor plot revealed a textbook case of heteroscedasticity. Performing a log transformation on our outcome variable, gross revenue, improved this issue. However, there is still a slight pattern noticeable in the residual vs. predictor plot. Analysis of the residual vs. quantitative predictor plots reveals the number of votes predictor as the likely source of this error. The caution to keep in mind with the inclusion of this predictor is the idea of participation bias. In further investigations, we intend to evaluate whether to "override" the LASSO results and exclude the number of votes predictor from the model if we determine that it is introducing bias into our model.

It is also important to consider any harms that may come from our analyses. Analyzing and predicting gross revenue of movies can have harmful effects on the performance of movies in box offices. For example, if we use our model to predict the success of a new movie, our model's prediction may affect people's decision as to whether or not to see the movie. This could affect the actual outcome of how movies perform at the box office by introducing bias into people's decision regarding whether to see a newly released movie.

# Update your models to use natural splines for some of the quantitative predictors to account for non-linearity (these are GAMs).

```{r}
# Build the GAM

gam_spec <- 
  gen_additive_mod() %>%
  set_engine(engine = 'mgcv') %>%
  set_mode('regression')

gam_mod <- fit(gam_spec,
    Gross ~ s(Runtime) + IMDB_Rating + s(Meta_score) + s(No_of_Votes) + Genre_1,
    data = imdb_clean 
)

```

```{r}
# Diagnostics: Check to see if the number of knots is large enough (if p-value is low, increase number of knots)

gam_mod %>% pluck('fit') %>% mgcv::gam.check() 
```

```{r}
gam_mod %>% pluck('fit') %>% summary() 
```

```{r}
# Visualize: Look at the estimated non-linear functions

gam_mod %>% pluck('fit') %>% plot()
```

```{r}
spline_rec <- recipe(Gross ~ ., data = imdb_clean) %>%
    step_rm(Series_Title, Released_Year, Overview,
            Director, Star1, Star2, Star3, Star4, Genre_2, Genre_3) %>% #variables causing issues
    step_nzv(all_predictors()) %>% 
    step_normalize(all_numeric_predictors()) %>% 
    step_dummy(all_nominal_predictors()) %>%
    step_naomit(Gross) %>%
    step_ns(Runtime, deg_free = 2) %>%
    step_ns(No_of_Votes, deg_free = 6)

spline_rec %>% prep(imdb_clean) %>% juice()
```

```{r}
# Use the edf (round to integer) from above to create a recipe by adding step_ns() for the variables you want model with a non-linear relationship and do CV. Compare this model to one without any splines.

lm_spec_gam <-
  linear_reg() %>%
  set_engine(engine = 'lm') %>%
  set_mode('regression')

spline_wf <- workflow() %>%
    add_model(lm_spec) %>%
    add_recipe(spline_rec)

cv_output <- fit_resamples( 
  spline_wf, # workflow
  resamples = data_cv10, # cv folds
  metrics = metric_set(mae,rmse,rsq)
)

cv_output %>% collect_metrics()

```

```{r}
new_mod <- spline_wf %>%
  fit(data = imdb_clean)

new_mod %>%
  tidy()

new_mod %>% 
  tidy() %>% 
  arrange(desc(abs(statistic)))

new_modcv <- fit_resamples(spline_wf, resamples = data_cv10, metrics = metric_set(mae, rmse, rsq))
```

# Compare insights from variable importance analyses here and the corresponding results from the Investigation

```{r}
# Lasso Model Spec with tune
gam_lasso_spec_tune <- 
  linear_reg() %>%
  set_args(mixture = 1, penalty = tune()) %>% ## mixture = 1 indicates Lasso
  set_engine(engine = 'glmnet') %>% #note we are using a different engine
  set_mode('regression')

# Recipe with standardization (!) --> just include all these always
data_rec <- recipe(Gross ~ ., data = imdb_clean) %>%
    step_rm(Series_Title, Released_Year, Overview,
            Director, Star1, Star2, Star3, Star4, Genre_2, Genre_3) %>%
    step_nzv(all_predictors()) %>% # removes variables with the same value (so duplicates don't mess up model)
    step_novel(all_nominal_predictors()) %>% # important if you have rare categorical variables 
    step_normalize(all_numeric_predictors()) %>%  # super important standardization step for LASSO
    step_dummy(all_nominal_predictors()) %>%  # creates indicator variables for categorical variables
    step_naomit(Gross)

# Workflow (Recipe + Model)
lasso_wf_tune1 <- workflow() %>% 
  add_recipe(data_rec) %>%
  add_model(gam_lasso_spec_tune)

# Tune Model (trying a variety of values of Lambda penalty)
penalty_grid <- grid_regular(
  penalty(range = c(-3, 1)),
  levels = 30)

tune_res1 <- tune_grid( # new function for tuning parameters
  lasso_wf_tune1, # workflow
  resamples = data_cv10, # cv folds
  metrics = metric_set(rmse, mae),
  grid = penalty_grid # penalty grid defined above
)

# Visualize Model Evaluation Metrics from Tuning
autoplot(tune_res1) + theme_classic()
```

# Collect CV Metrics and Select Best Model

```{r}

# Summarize Model Evaluation Metrics (CV)
collect_metrics(tune_res) %>%
  filter(.metric == 'rmse') %>% # or choose mae
  select(penalty, rmse = mean)

best_penalty1 <- select_best(tune_res1, metric = 'rmse') # choose penalty value based on lowest mae or rmse
```

# Fit Final Model

```{r}

# Fit Final Model
final_wf1 <- finalize_workflow(lasso_wf_tune1, best_penalty) # incorporates penalty value to workflow

final_fit1 <- fit(final_wf1, data = imdb_clean)

tidy(final_fit1)
```

```{r}
# Final ("best") model predictors and coefficients

final_fit1 %>% tidy() %>% filter(estimate != 0)
```

# Visualize Residuals

```{r}

# Visualize Residuals
lasso_mod_out1 <- new_mod %>%
    predict(new_data = imdb_clean) %>%
    bind_cols(imdb_clean) %>%
    mutate(resid = Gross - .pred)

ggplot(lasso_mod_out1, aes(x = .pred, y = resid)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") + 
    theme_classic()

ggplot(lasso_mod_out1, aes(x = Runtime, y = resid)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") + 
    theme_classic()

ggplot(lasso_mod_out1, aes(x = IMDB_Rating, y = resid)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") + 
    theme_classic()

ggplot(lasso_mod_out1, aes(x = No_of_Votes, y = resid)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") + 
    theme_classic()
```

# Compare model performance between your GAM models that the models that assuming linearity.

```{r}
full_modelcv %>% collect_metrics() #OLS
```

```{r}
lasso_fit %>% collect_metrics() #LASSO
```

```{r}
new_modcv %>% collect_metrics() #GAM
```

# Summarize investigations

Decide on an overall best model based on your investigations so far. To do this, make clear your analysis goals. Predictive accuracy? Interpretability? A combination of both?

We want a combination of both, our spline model has the lowest RMSE but could be difficult to interpret. We need more visualization and analysis before we move forward. However we are confident in moving forward.

# Societal Impact

Being that the data here is about movies, the harm that could come from our analysis would be minimal. 
Representation of different races, genders, abilities, etc. is an issue in Hollywood/the greater film industry, and it's likely that the harm our model/analysis produces is already in line with the issues that created and maintain this.
This is because the data we have is on past movies that have come out of this system. This means that whatever has made a movie successful, whether it's due to certain actors or directors, or whatever — this is a standard that has already existed. If we see the movies that have done really well because the cast is predominantly white, well this has been going on a for a long time because Hollywood is racist. Our model can't account for that, but we are aware of it.