### 2022-11-01 Logistic_Regression_Model.Rmd
Started Logistic Regression Model

---
title: 'STAT 253 Final Project - Classification Models'
author: "Jacob, Julian, Aristo, and Emily"
date: "October 31, 2022"
output: 
  html_document:
    df_print: paged
    code_download: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE, message=FALSE, warning=FALSE)
```

# Library Statements

```{r}
library(ISLR)
library(dplyr)
library(readr)
library(broom)
library(ggplot2)
library(tidymodels) 
library(stringr)
library(splitstackshape)
library(tidymodels)
library(lubridate)
library(rpart.plot)
library(cluster)
library(forcats)
tidymodels_prefer()
library(probably) #install.packages('probably')
library(vip)
```

# Read in Data

```{r}
library(readr)
imdb_top_1000 <- read_csv("~/Desktop/Statistical Machine Learning/R Files/Final Project/imdb_top_1000.csv")
```

# Data Cleaning

```{r}
imdb_clean <- imdb_top_1000 %>%
  select(-Poster_Link, -Certificate) %>%
  cSplit("Genre", sep = ",", direction = "wide") %>%
  mutate(Gross = log(Gross))

runtime_clean <- imdb_top_1000$Runtime %>%
  str_replace(" min", "") %>%
  as.numeric()

imdb_clean$Runtime <- runtime_clean
```
# Logistic Model with LASSO

```{r}
set.seed(123)

# Make sure you set reference level (to the outcome you are NOT interested in)
spam <- spam %>%
  mutate(spam = relevel(factor(spam), ref='not_spam')) #set reference level

data_cv10 <- vfold_cv(spam, v = 10)

# Logistic LASSO Regression Model Spec
logistic_lasso_spec_tune <- logistic_reg() %>%
    set_engine('glmnet') %>%
    set_args(mixture = 1, penalty = tune()) %>%
    set_mode('classification')

# Recipe
logistic_rec <- recipe(_____, data = spam) %>%
    step_normalize(all_numeric_predictors()) %>% 
    step_dummy(all_nominal_predictors())

# Workflow (Recipe + Model)
log_lasso_wf <- workflow() %>% 
    add_recipe(logistic_rec) %>%
    add_model(logistic_lasso_spec_tune)

# Tune Model (trying a variety of values of Lambda penalty)
penalty_grid <- grid_regular(
  penalty(range = c(__,__)), #log10 transformed 
  levels = 100)

tune_output <- tune_grid( 
  log_lasso_wf, # workflow
  resamples = data_cv10, # cv folds
  metrics = metric_set(roc_auc,accuracy),
  control = control_resamples(save_pred = TRUE, event_level = 'second'),
  grid = penalty_grid # penalty grid defined above
)

# Visualize Model Evaluation Metrics from Tuning
autoplot(tune_output) + theme_classic()
```

### Inspecting the Model

```{r}
# Visualize Model Evaluation Metrics from Tuning
autoplot(tune_output) + theme_classic()
```

```{r}
# Select Penalty
best_se_penalty <- select_by_one_std_err(tune_output, metric = 'roc_auc', desc(penalty)) # choose penalty value based on the largest penalty within 1 se of the highest CV roc_auc
best_se_penalty

# Fit Final Model
final_fit_se <- finalize_workflow(log_lasso_wf, best_se_penalty) %>% # incorporates penalty value to workflow 
    fit(data = spam)

final_fit_se %>% tidy()
```

```{r}
# Comment on Variable Importance

glmnet_output <- final_fit_se %>% extract_fit_engine()

# Create a boolean matrix (predictors x lambdas) of variable exclusion
bool_predictor_exclude <- glmnet_output$beta==0

# Loop over each variable
var_imp <- sapply(seq_len(nrow(bool_predictor_exclude)), function(row) {
    this_coeff_path <- bool_predictor_exclude[row,]
    if(sum(this_coeff_path) == ncol(bool_predictor_exclude)){ return(0)}else{
    return(ncol(bool_predictor_exclude) - which.min(this_coeff_path) + 1)}
})

# Create a dataset of this information and sort
var_imp_data <- tibble(
    var_name = rownames(bool_predictor_exclude),
    var_imp = var_imp
)
var_imp_data %>% arrange(desc(var_imp))
```

# Evaluation Metrics

```{r}

# CV results for "best lambda"
tune_output %>%
    collect_metrics() %>%
    filter(penalty == best_se_penalty %>% pull(penalty))

# Count up number of spam and not_spam emails in the training data
spam %>%
    count(spam) # Name of the outcome variable goes inside count()

# Compute the NIR
```

### Final Model

```{r}
# Soft Predictions on Training Data
final_output <- final_fit_se %>% predict(new_data = spam, type='prob') %>% bind_cols(spam)
```

```{r}
# Use soft predictions
final_output %>%
    roc_curve(spam,.pred_spam,event_level = 'second') %>%
    autoplot()
```

```{r}
# thresholds in terms of reference level
threshold_output <- final_output %>%
    threshold_perf(truth = spam, estimate = .pred_not_spam, thresholds = seq(0,1,by=.01))

# J-index v. threshold for not_spam
threshold_output %>%
    filter(.metric == 'j_index') %>%
    ggplot(aes(x = .threshold, y = .estimate)) +
    geom_line() +
    labs(y = 'J-index', x = 'threshold') +
    theme_classic()

threshold_output %>%
    filter(.metric == 'j_index') %>%
    arrange(desc(.estimate))

# Distance v. threshold for not_spam

threshold_output %>%
    filter(.metric == 'distance') %>%
    ggplot(aes(x = .threshold, y = .estimate)) +
    geom_line() +
    labs(y = 'Distance', x = 'threshold') +
    theme_classic()

threshold_output %>%
    filter(.metric == 'distance') %>%
    arrange(.estimate)

log_metrics <- metric_set(accuracy,sens,yardstick::spec)

final_output %>%
    mutate(.pred_class = make_two_class_pred(.pred_not_spam, levels(spam), threshold = ___ )) %>%
    log_metrics(truth = spam, estimate = .pred_class, event_level = 'second')
```