### 2022-11-28 ClusteringModel.Rmd
Started Clustering (K-Means)

---
title: "Clustering Model"
author: "Emily Bergin"
date: '2022-11-28'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE, message=FALSE, warning=FALSE)
```

# Library Statements

```{r}
library(dplyr)
library(purrr)
library(ggplot2)
library(ISLR)
library(readr)
library(tidymodels)
library(stringr)
library(lubridate)
```

# Read in Data

```{r}
imdb_top_1000 <- read_csv("~/Desktop/Statistical Machine Learning/R Files/Final Project/imdb_top_1000_CLEAN.csv")
```

# Data Cleaning

```{r}
imdb_clean <- imdb_top_1000 %>%
  cSplit("Genre", sep = ",", direction = "wide") %>%
  mutate(Gross = log(Revenue-Budget)) %>%
  select(-...15)

runtime_clean <- imdb_top_1000$Runtime %>%
  str_replace(" min", "") %>%
  as.numeric()

imdb_clean$Runtime <- runtime_clean

imdb_clean <- imdb_clean %>%
  filter(Gross != "-Inf") %>%
  filter(!is.na(Gross)) %>%
  filter(!is.na(Budget))
```

### K-Means Clustering

```{r}
# Select just the bill length and depth variables
imdb_sub <- imdb_clean %>%
    select(Budget, Gross)

# Run k-means for k = centers = 3
set.seed(253)
kclust_k10 <- kmeans(scale(imdb_sub), centers = 10)

# Display the cluter assignments
kclust_k10$cluster

# Add a variable (kclust_3) to the original dataset 
# containing the cluster assignments
imdb_clean <- imdb_clean %>%
    mutate(kclust_10 = factor(kclust_k10$cluster))
```

```{r}
# Visualize the cluster assignments on the original scatterplot
imdb_clean %>%
  ggplot(aes(x = Budget, y = Gross, color = kclust_10)) +
    geom_point() + theme_classic()
```

## Interpreting the Clusters

```{r}
imdb_clean %>%
    group_by(kclust_10) %>%
    summarize(across(c(Gross, Budget), mean))
```