### 2022-11-29 ClusteringModel.Rmd
Worked on K-Means and started Hierarchial Clustering

---
title: "Clustering Model"
author: "Emily Bergin"
date: '2022-11-28'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE, message=FALSE, warning=FALSE)
```

# Library Statements

```{r}
library(dplyr)
library(purrr)
library(ggplot2)
library(ISLR)
library(readr)
library(tidymodels)
library(stringr)
library(lubridate)
library(splitstackshape)
```

# Read in Data

```{r}
imdb_top_1000 <- read_csv("~/Desktop/Statistical Machine Learning/R Files/Final Project/imdb_top_1000_CLEAN.csv")
```

# Data Cleaning

```{r}
imdb_clean <- imdb_top_1000 %>%
  cSplit("Genre", sep = ",", direction = "wide") %>%
  mutate(Gross = log(Revenue-Budget)) %>%
  select(-...15)

runtime_clean <- imdb_top_1000$Runtime %>%
  str_replace(" min", "") %>%
  as.numeric()

imdb_clean$Runtime <- runtime_clean

imdb_clean <- imdb_clean %>%
  filter(Gross != "-Inf") %>%
  filter(!is.na(Gross)) %>%
  filter(!is.na(Budget))
```

### K-Means Clustering

```{r}
# Select just the bill length and depth variables
imdb_sub <- imdb_clean %>%
    select(Budget, Gross)

# Run k-means for k = centers = 3
set.seed(253)
kclust_k10 <- kmeans(scale(imdb_sub), centers = 10)

# Display the cluter assignments
kclust_k10$cluster

# Add a variable (kclust_3) to the original dataset 
# containing the cluster assignments
imdb_clean <- imdb_clean %>%
    mutate(kclust_10 = factor(kclust_k10$cluster))
```

```{r}
# Visualize the cluster assignments on the original scatterplot
imdb_clean %>%
  ggplot(aes(x = Budget, y = Gross, color = kclust_10)) +
    geom_point() + theme_classic()
```

```{r}
# Data-specific function to cluster and calculate total within-cluster SS
imdb_cluster_ss <- function(k){
    # Perform clustering
    kclust <- kmeans(scale(imdb_sub), centers = k)

    # Return the total within-cluster sum of squares
    return(kclust$tot.withinss)
}

tibble(
    k = 1:15,
    tot_wc_ss = purrr::map_dbl(1:15, imdb_cluster_ss)
) %>% 
    ggplot(aes(x = k, y = tot_wc_ss)) +
    geom_point() + 
    labs(x = "Number of clusters",y = 'Total within-cluster sum of squares') + 
    theme_classic()
```

```{r}
kclust_k8 <- kmeans(scale(imdb_sub), centers = 8)

# Display the cluter assignments
kclust_k8$cluster

# Add a variable (kclust_3) to the original dataset 
# containing the cluster assignments
imdb_clean <- imdb_clean %>%
    mutate(kclust_8 = factor(kclust_k8$cluster))
```

```{r}
# Visualize the cluster assignments on the original scatterplot
imdb_clean %>%
  ggplot(aes(x = Budget, y = Gross, color = kclust_8)) +
    geom_point() + theme_classic()
```

*Pick either k = 8 or k = 10 --> there are 7 genres with larger n and 13 total genres. Film noir is fancy way of saying crime drama (ex. Knives Out). One thing to note is a significant number of movies have Genre_2 listed as Family Mystery and Horror (and many have one of the main genres listed as genre 2 as well). What is a good way to summarize both Genre 1 and 2?*

```{r}
imdb_clean %>%
  count(Genre_1)

imdb_clean %>%
  count(Genre_2)
```

## Interpreting the Clusters

```{r}
imdb_clean %>%
  group_by(kclust_10) %>%
  summarize(across(c(Gross, Budget), mean))

imdb_clean %>%
  group_by(kclust_8) %>%
  count(Genre_1)

imdb_clean %>%
  group_by(kclust_8) %>%
  count(Genre_2)

imdb_clean%>%
  count(kclust_8)
```

### Hierarchial Clustering

```{r}
# Random subsample of 50 penguins
set.seed(253)
imdb_hc <- imdb_clean %>%
    slice_sample(n = 25)

# Select the variables to be used in clustering
imdb_hc_sub <- imdb_hc %>%
    select(Gross, Budget)

# Summary statistics for the variables
summary(imdb_hc_sub)

# Compute a distance matrix on the scaled data
dist_mat_scaled <- dist(scale(imdb_hc_sub))
```

```{r}
imdb_hc_avg <- hclust(dist_mat_scaled, method = "average")

# Plot dendrogram
plot(imdb_hc_avg)
```

```{r}

plot(imdb_hc_avg, labels = imdb_hc$Genre_1)

plot(imdb_hc_avg, labels = paste(imdb_hc$Genre_1, imdb_hc$Genre_2))
```