### 2022-11-29 Final_Project_Code_ALL.Rmd
Worked on classification models

---
title: "Final Project Code ALL"
author: "Emily, Julian, Jacob, and Aristo"
date: '2022-11-20'
output: 
  html_document:
    df_print: paged
    code_download: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE, message=FALSE, warning=FALSE)
```

# Library Statements

```{r}
library(ISLR)
library(dplyr)
library(readr)
library(broom)
library(ggplot2)
library(tidymodels) 
library(stringr)
library(splitstackshape)
library(lubridate)
library(rpart.plot)
library(cluster)
library(forcats)
tidymodels_prefer()
library(probably) #install.packages('probably')
library(vip)
```

# Read in Data

```{r}
imdb_top_1000 <- read_csv("~/Desktop/Statistical Machine Learning/R Files/Final Project/imdb_top_1000_CLEAN.csv")
```

# Data Cleaning

```{r}
imdb_clean <- imdb_top_1000 %>%
  cSplit("Genre", sep = ",", direction = "wide") %>%
  mutate(Gross = log(Revenue-Budget)) %>%
  select(-...15)

runtime_clean <- imdb_top_1000$Runtime %>%
  str_replace(" min", "") %>%
  as.numeric()

imdb_clean$Runtime <- runtime_clean

imdb_clean <- imdb_clean %>%
  filter(Gross != "-Inf") %>%
  drop_na(Gross, Budget)
```

# Creation of CV Folds

```{r}
data_cv10 <- vfold_cv(imdb_clean, v = 10)
```

# Model Spec for Linear Regression Model

```{r}
lm_spec <-
    linear_reg() %>% 
    set_engine(engine = 'lm') %>% 
    set_mode('regression')
```

# Recipes and Workflows

```{r}
full_rec <- recipe(Gross ~ Runtime + IMDB_Rating + Meta_score + 
                   No_of_Votes + Genre_1, data = imdb_clean) %>%
    step_nzv(all_predictors()) %>% 
    step_normalize(all_numeric_predictors()) %>% 
    step_dummy(all_nominal_predictors()) %>%
    step_naomit(Gross)

full_lm_wf <- workflow() %>%
    add_recipe(full_rec) %>%
    add_model(lm_spec)

full_model <- fit(full_lm_wf, data = imdb_clean)

full_modelcv <- fit_resamples(full_lm_wf, resamples = data_cv10, metrics = metric_set(rmse, rsq, mae))

full_model %>% tidy()
```

# LASSO - Fit and Tune Models

```{r}

# Lasso Model Spec with tune
lm_lasso_spec_tune <- 
  linear_reg() %>%
  set_args(mixture = 1, penalty = tune()) %>% ## mixture = 1 indicates Lasso
  set_engine(engine = 'glmnet') %>% #note we are using a different engine
  set_mode('regression')

# Recipe with standardization (!) --> just include all these always
data_rec <- recipe(Gross ~ Runtime + IMDB_Rating + Meta_score + 
                   No_of_Votes + Genre_1, data = imdb_clean) %>%
    step_nzv(all_predictors()) %>% # removes variables with the same value (so duplicates don't mess up model)
    step_novel(all_nominal_predictors()) %>% # important if you have rare categorical variables 
    step_normalize(all_numeric_predictors()) %>%  # super important standardization step for LASSO
    step_dummy(all_nominal_predictors()) %>%  # creates indicator variables for categorical variables
    step_naomit(Gross)

# Workflow (Recipe + Model)
lasso_wf_tune <- workflow() %>% 
  add_recipe(data_rec) %>%
  add_model(lm_lasso_spec_tune)

# Tune Model (trying a variety of values of Lambda penalty)
penalty_grid <- grid_regular(
  penalty(range = c(-3, 1)),
  levels = 30)

tune_res <- tune_grid( # new function for tuning parameters
  lasso_wf_tune, # workflow
  resamples = data_cv10, # cv folds
  metrics = metric_set(rmse, mae),
  grid = penalty_grid # penalty grid defined above
)

# Visualize Model Evaluation Metrics from Tuning
autoplot(tune_res) + theme_classic()
```

# Collect CV Metrics and Select Best Model

```{r}

# Summarize Model Evaluation Metrics (CV)
lasso_mod <- collect_metrics(tune_res) %>%
  filter(.metric == 'rmse') %>% # or choose mae
  select(penalty, rmse = mean)

best_penalty <- select_best(tune_res, metric = 'rmse') # choose penalty value based on lowest mae or rmse

lasso_mod
```

# Fit Final Model

```{r}

# Fit Final Model
final_wf <- finalize_workflow(lasso_wf_tune, best_penalty) # incorporates penalty value to workflow

final_fit <- fit(final_wf, data = imdb_clean)

lasso_fit <- fit_resamples(final_wf, resamples = data_cv10, metrics = metric_set(rmse, rsq, mae))

tidy(final_fit)
```

```{r}
# Final ("best") model predictors and coefficients

final_fit %>% tidy() %>% filter(estimate != 0)
```

# Visualize Residuals

```{r}

# Visualize Residuals
lasso_mod_out <- final_fit %>%
    predict(new_data = imdb_clean) %>%
    bind_cols(imdb_clean) %>%
    mutate(resid = Gross - .pred)

ggplot(lasso_mod_out, aes(x = .pred, y = resid)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") + 
    theme_classic()

ggplot(lasso_mod_out, aes(x = Runtime, y = resid)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") + 
    theme_classic()

ggplot(lasso_mod_out, aes(x = IMDB_Rating, y = resid)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") + 
    theme_classic()

ggplot(lasso_mod_out, aes(x = No_of_Votes, y = resid)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") + 
    theme_classic()
```

The overall goal of our linear regression model is to predict a movie's gross box office revenue based on a number of predictors. The goal of our initial investigation was to determine which subset of predictors are the most important for inclusion in our model. In order to determine the "best" linear regression model to achieve our analysis goals, we utilized the LASSO algorithm. The goal of LASSO is to decrease variance in our model by constraining the coefficient estimates and discouraging the inclusion of weakly informative predictors through the use of a tuning parameter. This improves the predictive accuracy of our model. We chose this algorithm over the best subset selection algorithm because LASSO is substantially more computationally efficient. Furthermore, since our main analysis goal is predictive accuracy, we decided against using forward stepwise or backwards stepwise selection because these algorithms do not guarantee the best overall model. A model using a smaller subset of variables may have increased interpretability, but this would come at the expense of predictive accuracy.

The main caution we want to keep in mind when communicating our work is that our model may be bias against new names in the movie industry. The model cannot determine a director, actor, or actresses' talent. It only bases its predictions off of how previous movies that individual was involved in previously performed. Furthermore, the inclusion of the number of votes predictor in our model raises further concerns of potential bias. Our initial evaluation of the residual vs. predictor plot revealed a textbook case of heteroscedasticity. Performing a log transformation on our outcome variable, gross revenue, improved this issue. However, there is still a slight pattern noticeable in the residual vs. predictor plot. Analysis of the residual vs. quantitative predictor plots reveals the number of votes predictor as the likely source of this error. The caution to keep in mind with the inclusion of this predictor is the idea of participation bias. In further investigations, we intend to evaluate whether to "override" the LASSO results and exclude the number of votes predictor from the model if we determine that it is introducing bias into our model.

It is also important to consider any harms that may come from our analyses. Analyzing and predicting gross revenue of movies can have harmful effects on the performance of movies in box offices. For example, if we use our model to predict the success of a new movie, our model's prediction may affect people's decision as to whether or not to see the movie. This could affect the actual outcome of how movies perform at the box office by introducing bias into people's decision regarding whether to see a newly released movie.

# GAMs Using Splines

```{r}
# Build the GAM

gam_spec <- 
  gen_additive_mod() %>%
  set_engine(engine = 'mgcv') %>%
  set_mode('regression')

gam_mod <- fit(gam_spec,
    Gross ~ s(Runtime) + IMDB_Rating + s(Meta_score) + s(No_of_Votes) + Genre_1,
    data = imdb_clean 
)

```

```{r}
# Diagnostics: Check to see if the number of knots is large enough (if p-value is low, increase number of knots)

gam_mod %>% pluck('fit') %>% mgcv::gam.check() 
```

```{r}
gam_mod %>% pluck('fit') %>% summary() 
```

```{r}
# Visualize: Look at the estimated non-linear functions

gam_mod %>% pluck('fit') %>% plot()
```

```{r}
spline_rec <- recipe(Gross ~ Runtime + IMDB_Rating + Meta_score + 
                   No_of_Votes + Genre_1, data = imdb_clean) %>%
    step_nzv(all_predictors()) %>% 
    step_normalize(all_numeric_predictors()) %>% 
    step_dummy(all_nominal_predictors()) %>%
    step_naomit(Gross) %>%
    step_ns(Runtime, deg_free = 2) %>%
    step_ns(No_of_Votes, deg_free = 6)

spline_rec %>% prep(imdb_clean) %>% juice()
```

```{r}
# Use the edf (round to integer) from above to create a recipe by adding step_ns() for the variables you want model with a non-linear relationship and do CV. Compare this model to one without any splines.

lm_spec_gam <-
  linear_reg() %>%
  set_engine(engine = 'lm') %>%
  set_mode('regression')

spline_wf <- workflow() %>%
    add_model(lm_spec) %>%
    add_recipe(spline_rec)

cv_output <- fit_resamples( 
  spline_wf, # workflow
  resamples = data_cv10, # cv folds
  metrics = metric_set(mae,rmse,rsq)
)

cv_output %>% collect_metrics()

```

```{r}
new_mod <- spline_wf %>%
  fit(data = imdb_clean)

new_mod %>%
  tidy()

new_mod %>% 
  tidy() %>% 
  arrange(desc(abs(statistic)))

new_modcv <- fit_resamples(spline_wf, resamples = data_cv10, metrics = metric_set(mae, rmse, rsq))
```

# Variable Importance for GAM

```{r}
# Lasso Model Spec with tune
gam_lasso_spec_tune <- 
  linear_reg() %>%
  set_args(mixture = 1, penalty = tune()) %>% ## mixture = 1 indicates Lasso
  set_engine(engine = 'glmnet') %>% #note we are using a different engine
  set_mode('regression')

# Recipe with standardization (!) --> just include all these always
data_rec <- recipe(Gross ~ Runtime + IMDB_Rating + Meta_score + 
                   No_of_Votes + Genre_1, data = imdb_clean) %>%
    step_nzv(all_predictors()) %>% # removes variables with the same value (so duplicates don't mess up model)
    step_novel(all_nominal_predictors()) %>% # important if you have rare categorical variables 
    step_normalize(all_numeric_predictors()) %>%  # super important standardization step for LASSO
    step_dummy(all_nominal_predictors()) %>%  # creates indicator variables for categorical variables
    step_naomit(Gross)

# Workflow (Recipe + Model)
lasso_wf_tune1 <- workflow() %>% 
  add_recipe(data_rec) %>%
  add_model(gam_lasso_spec_tune)

# Tune Model (trying a variety of values of Lambda penalty)
penalty_grid <- grid_regular(
  penalty(range = c(-3, 1)),
  levels = 30)

tune_res1 <- tune_grid( # new function for tuning parameters
  lasso_wf_tune1, # workflow
  resamples = data_cv10, # cv folds
  metrics = metric_set(rmse, mae),
  grid = penalty_grid # penalty grid defined above
)

# Visualize Model Evaluation Metrics from Tuning
autoplot(tune_res1) + theme_classic()
```

# Collect CV Metrics and Select Best Model

```{r}

# Summarize Model Evaluation Metrics (CV)
collect_metrics(tune_res) %>%
  filter(.metric == 'rmse') %>% # or choose mae
  select(penalty, rmse = mean)

best_penalty1 <- select_best(tune_res1, metric = 'rmse') # choose penalty value based on lowest mae or rmse
```

# Fit Final Model

```{r}

# Fit Final Model
final_wf1 <- finalize_workflow(lasso_wf_tune1, best_penalty) # incorporates penalty value to workflow

final_fit1 <- fit(final_wf1, data = imdb_clean)

tidy(final_fit1)
```

```{r}
# Final ("best") model predictors and coefficients

final_fit1 %>% tidy() %>% filter(estimate != 0)
```

# Visualize Residuals

```{r}

# Visualize Residuals
lasso_mod_out1 <- new_mod %>%
    predict(new_data = imdb_clean) %>%
    bind_cols(imdb_clean) %>%
    mutate(resid = Gross - .pred)

ggplot(lasso_mod_out1, aes(x = .pred, y = resid)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") + 
    theme_classic()

ggplot(lasso_mod_out1, aes(x = Runtime, y = resid)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") + 
    theme_classic()

ggplot(lasso_mod_out1, aes(x = IMDB_Rating, y = resid)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") + 
    theme_classic()

ggplot(lasso_mod_out1, aes(x = No_of_Votes, y = resid)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") + 
    theme_classic()
```

# Comparing Model Performance

```{r}
full_modelcv %>% collect_metrics() #OLS
```

```{r}
lasso_fit %>% collect_metrics() #LASSO
```

```{r}
new_modcv %>% collect_metrics() #GAM
```

# Summarize Investigations

The goal of our investigations is to accurately predict a movie's gross box office gross revenue based on a number of predictors. We compared three different model types: ordinary linear regression (OLS) with all predictors, OLS using LASSO to select a subset of variables, and a GAM using splines to account for nonlinearity. The GAM model has the lowest RMSE but could be difficult to interpret. We are looking forward to continuing our investigations using Decision Trees which may offer a more easily interpreted method for modeling nonlinearity. Ideally, we would like an accurate model that is relatively easy to interpret.

# Societal Impact

In addition to our earlier discussion of societal impacts that our model may have, we want to keep in mind the biases of the Hollywood movie industry. Representation of different races, genders, abilities, etc. is an issue in Hollywood/the greater film industry, and it is likely that the harm our model/analysis produces is already in line with the issues that created and maintain this. This is because the data we have is from past movies that came out of this system. This means that whatever made a movie successful, whether it's due to certain actors, directors, etc., is a standard that already existed. If we see the movies that have done really well have predominantly white casts, then it is likely a result of a longstanding issue of racism in the movie industry. Our model can't account for that, but we are aware of it.

```{r}
imdb_clean_class <- imdb_clean %>%
  mutate(success_ratio = Revenue/Budget) %>%
  mutate(flop = as.factor(ifelse(success_ratio > 2, 'FALSE', 'TRUE'))) %>%
  drop_na(flop, No_of_Votes,Runtime, IMDB_Rating,Meta_score,Genre_1)
  # filter(!is.na(flop)) %>%
  # filter(Gross != "NaN") %>%
  # filter(success_ratio != "Inf") %>%
  # filter(Budget > 10)
```

# Bagging and Random Forests

```{r}
# Model Specification
rf_spec <- rand_forest() %>%
  set_engine(engine = 'ranger') %>% 
  set_args(trees = 1000, # Number of trees
           min_n = NULL,
           probability = FALSE, # FALSE: get hard predictions (not needed for regression)
           importance = 'impurity') %>% # we'll come back to this at the end
  set_mode('classification') # change this for regression

# Recipe
data_rec <- recipe(flop ~ No_of_Votes + Runtime + IMDB_Rating + Meta_score + Genre_1,
                   data = imdb_clean_class) %>%
  step_naomit(flop, No_of_Votes, Runtime, IMDB_Rating, Meta_score, Genre_1)

# Workflows
data_wf_mtry3 <- workflow() %>%
  add_model(rf_spec %>% set_args(mtry = 3)) %>%
  add_recipe(data_rec)

data_wf_mtry4 <- workflow() %>%
  add_model(rf_spec %>% set_args(mtry = 4)) %>%
  add_recipe(data_rec)

## Create workflows for min_n = 10, 25, and 50

data_wf_mtry5 <- workflow() %>%
  add_model(rf_spec %>% set_args(mtry = 5)) %>%
  add_recipe(data_rec)
```

```{r}
# Fit Models
set.seed(123) # make sure to run this before each fit so that you have the same 1000 trees
data_fit_mtry3 <- fit(data_wf_mtry3, data = imdb_clean_class)

set.seed(123) 
data_fit_mtry4 <- fit(data_wf_mtry4, data = imdb_clean_class)

set.seed(123)
data_fit_mtry5 <- fit(data_wf_mtry5, data = imdb_clean_class)
```

```{r}
# Custom Function to get OOB predictions, true observed outcomes and add a user-provided model label
rf_OOB_output <- function(fit_model, model_label, truth){
    tibble(
          .pred_class = fit_model %>% extract_fit_engine() %>% pluck('predictions'), #OOB predictions
          flop = truth,
          label = model_label
      )
}
```

```{r}
# Evaluate OOB Metrics

data_rf_OOB_output <- bind_rows(
    rf_OOB_output(data_fit_mtry3,3, imdb_clean_class %>% pull(flop)),
    rf_OOB_output(data_fit_mtry4,4, imdb_clean_class %>% pull(flop)),
    rf_OOB_output(data_fit_mtry5,5, imdb_clean_class %>% pull(flop))
)

data_rf_OOB_output %>% 
    group_by(label) %>%
    accuracy(truth = flop, estimate = .pred_class)
```

### Evaluating the Forest

```{r}
rf_OOB_output(data_fit_mtry4,4, imdb_clean_class %>% pull(flop)) %>%
    conf_mat(truth = flop, estimate= .pred_class)
```

### Variable Importance

```{r}
# Impurity

model_output <-data_fit_mtry4 %>% 
    extract_fit_engine()

model_output %>% 
    vip(num_features = 10) + theme_classic() #based on impurity, 10 meaning the top 10

model_output %>% vip::vi() %>% head()
model_output %>% vip::vi() %>% tail()
```

```{r}
# Permutation

model_output2 <- data_wf_mtry4 %>% 
  update_model(rf_spec %>% set_args(importance = "permutation")) %>% #based on permutation
  fit(data = imdb_clean_class) %>% 
    extract_fit_engine()

model_output2 %>% 
    vip(num_features = 10) + theme_classic()

model_output2 %>% vip::vi() %>% head()
model_output2 %>% vip::vi() %>% tail()
```

# Logistic Regression

```{r}
set.seed(123)

# Logistic Regression Model Spec
logistic_spec <- logistic_reg() %>%
    set_engine('glm') %>%
    set_mode('classification')

# Recipe
logistic_rec <- recipe(flop ~ No_of_Votes + Runtime + IMDB_Rating + Genre_1,
                   data = imdb_clean_class)

# Workflow (Recipe + Model)
log_wf <- workflow() %>%
    add_recipe(logistic_rec) %>%
    add_model(logistic_spec)

# Fit Model
log_fit <- fit(log_wf, data = imdb_clean_class)

tidy(log_fit)
```

```{r}
log_fit %>% tidy() %>%
  mutate(OR = exp(estimate))
```

```{r}
predict(log_fit, new_data = data.frame(No_of_Votes = 10000, Runtime = 112, IMDB_Rating = 9.8,
                                        Genre_1 = "Drama"), type = "prob"
)
```

```{r}
predict(log_fit, new_data = data.frame(No_of_Votes = 10000, Runtime = 112, IMDB_Rating = 9.8,
                                        Genre_1 = "Drama"), type = "class"
)
```