### 2022-12-02 Draft_Final_Code.Rmd
worked on linear regression more

---
title: "Final Project Code ALL"
author: "Emily, Julian, Jacob, and Aristo"
date: '2022-11-20'
output: 
  html_document:
    df_print: paged
    toc: true
    code_download: true
    theme: paper
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE, message=FALSE, warning=FALSE)
```

# Library Statements

```{r}
library(ISLR)
library(dplyr)
library(readr)
library(broom)
library(ggplot2)
library(tidymodels) 
library(stringr)
library(splitstackshape)
library(lubridate)
library(rpart.plot)
library(cluster)
library(forcats)
tidymodels_prefer()
library(probably) #install.packages('probably')
library(vip)
```

# Dataset

```{r}
imdb_top_1000 <- read_csv("~/Desktop/Statistical Machine Learning/R Files/Final Project/imdb_top_1000_CLEAN.csv")
```

## Data Cleaning

```{r}
imdb_clean <- imdb_top_1000 %>%
  cSplit("Genre", sep = ",", direction = "wide") %>%
  mutate(Gross = log(Revenue-Budget)) %>%
  select(-...15)

runtime_clean <- imdb_top_1000$Runtime %>%
  str_replace(" min", "") %>%
  as.numeric()

imdb_clean$Runtime <- runtime_clean

imdb_clean <- imdb_clean %>%
  filter(Gross != "-Inf") %>%
  drop_na(Gross, Budget)
```

## Data

```{r}
head(imdb_clean)
```

# Regression Models

## Linear Regression Model

### Creation of CV Folds

```{r}
data_cv10 <- vfold_cv(imdb_clean, v = 10)
```

### Model Spec, Recipes, and Workflows for Linear Regression Model

```{r}
# Model Spec

lm_spec <-
    linear_reg() %>% 
    set_engine(engine = 'lm') %>% 
    set_mode('regression')

# Recipe

full_lm_rec <- recipe(Gross ~ Runtime + IMDB_Rating + Meta_score + 
                   No_of_Votes + Genre_1, data = imdb_clean) %>%
    step_nzv(all_predictors()) %>% 
    step_normalize(all_numeric_predictors()) %>% 
    step_dummy(all_nominal_predictors()) %>%
    step_naomit(Gross, Runtime, IMDB_Rating, Meta_score, No_of_Votes)

# Workflow

full_lm_wf <- workflow() %>%
    add_recipe(full_lm_rec) %>%
    add_model(lm_spec)
```

### Fit Preliminary Model

```{r}
full_lm_model <- fit(full_lm_wf, data = imdb_clean)

full_lm_model %>% tidy()
```

### Obtain Evaluation Metrics for Preliminary Model

```{r}
full_lm_modelcv <- fit_resamples(full_lm_wf, resamples = data_cv10, metrics = metric_set(rmse, rsq, mae))

full_lm_modelcv %>%
  collect_metrics()
```

### Perform LASSO for Subset Selection

```{r}
# Lasso Model Spec with tune

lm_lasso_spec_tune <- 
  linear_reg() %>%
  set_args(mixture = 1, penalty = tune()) %>%   # mixture = 1 indicates Lasso
  set_engine(engine = 'glmnet') %>%             
  set_mode('regression')

# Recipe

data_rec_lasso <- recipe(Gross ~ Runtime + IMDB_Rating + Meta_score + 
                   No_of_Votes + Genre_1, data = imdb_clean) %>%
    step_nzv(all_predictors()) %>%                # removes variables with the same value (don't want duplicates)
    step_novel(all_nominal_predictors()) %>%      # important if you have rare categorical variables 
    step_normalize(all_numeric_predictors()) %>%  # standardization important step for LASSO
    step_dummy(all_nominal_predictors()) %>%      # creates indicator variables for categorical variables
    step_naomit(Gross, Runtime, IMDB_Rating,      # omit any NA values
                Meta_score, No_of_Votes)

# Workflow

lasso_wf_tune <- workflow() %>% 
  add_recipe(data_rec_lasso) %>%
  add_model(lm_lasso_spec_tune)

# Tune Model (trying a variety of values of Lambda penalty)

penalty_grid <- grid_regular(
  penalty(range = c(-3, 1)),
  levels = 30)

tune_res <- tune_grid(
  lasso_wf_tune, 
  resamples = data_cv10, 
  metrics = metric_set(rmse, mae),
  grid = penalty_grid 
)

# Visualize Model Evaluation Metrics from Tuning

autoplot(tune_res) + theme_classic()

# Collect CV Metrics and Select Best Model

# Summarize Model Evaluation Metrics (CV)
lasso_mod <- collect_metrics(tune_res) %>%
  filter(.metric == 'rmse') %>%
  select(penalty, rmse = mean)

best_penalty <- select_best(tune_res, metric = 'rmse') # choose penalty value

lasso_mod
```

### Fit Final Model

```{r}
# Fit Final Model

final_wf <- finalize_workflow(lasso_wf_tune, best_penalty) # incorporates penalty value to workflow

final_fit <- fit(final_wf, data = imdb_clean)

lasso_fit <- fit_resamples(final_wf, resamples = data_cv10, metrics = metric_set(rmse, rsq, mae))

tidy(final_fit)

# Final ("best") model predictors and coefficients

final_fit %>% tidy() %>% filter(estimate != 0)
```

### Visualize Residuals

```{r}
lasso_mod_out <- final_fit %>%
    predict(new_data = imdb_clean) %>%
    bind_cols(imdb_clean) %>%
    mutate(resid = Gross - .pred)

ggplot(lasso_mod_out, aes(x = .pred, y = resid)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") + 
    theme_classic()

ggplot(lasso_mod_out, aes(x = Runtime, y = resid)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") + 
    theme_classic()

ggplot(lasso_mod_out, aes(x = IMDB_Rating, y = resid)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") + 
    theme_classic()

ggplot(lasso_mod_out, aes(x = No_of_Votes, y = resid)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") + 
    theme_classic()
```

## GAM Using Splines

```{r}
# Build the GAM

gam_spec <- 
  gen_additive_mod() %>%
  set_engine(engine = 'mgcv') %>%
  set_mode('regression')

gam_mod <- fit(gam_spec,
    Gross ~ s(Runtime) + IMDB_Rating + s(Meta_score) + s(No_of_Votes) + Genre_1,
    data = imdb_clean 
)

```

````{r}
# Diagnostics: Check to see if the number of knots is large enough (if p-value is low, increase number of knots)

gam_mod %>% pluck('fit') %>% mgcv::gam.check()

gam_mod %>% pluck('fit') %>% summary() 
```

```{r}
# Visualize: Look at the estimated non-linear functions

gam_mod %>% pluck('fit') %>% plot()
```