### 2022-12-12 Project_Final_Code.Rmd
Final Code

---
title: "Final Project Code"
author: "Emily, Julian, Jacob, and Aristo"
date: '2022-11-20'
output: 
  html_document:
    df_print: paged
    toc: true
    code_download: true
    theme: paper
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, tidy = TRUE)
```

# Library Statements

```{r}
library(ISLR)
library(dplyr)
library(readr)
library(broom)
library(ggplot2)
library(tidymodels) 
library(stringr)
library(splitstackshape)
library(lubridate)
library(rpart.plot)
library(cluster)
library(forcats)
tidymodels_prefer()
library(probably) #install.packages('probably')
library(vip)
```

# Dataset

```{r}
imdb_top_1000 <- read_csv("~/Desktop/Statistical Machine Learning/R Files/Final Project/imdb_top_1000_CLEAN.csv")
```

## Data Cleaning

```{r}
imdb_clean <- imdb_top_1000 %>%
  cSplit("Genre", sep = ",", direction = "wide") %>%
  mutate(Gross = log(Revenue-Budget))

runtime_clean <- imdb_top_1000$Runtime %>%
  str_replace(" min", "") %>%
  as.numeric()

imdb_clean$Runtime <- runtime_clean

imdb_clean <- imdb_clean %>%
  filter(Gross != "-Inf") %>%
  drop_na(Gross, Budget)
```

## Data

```{r}
head(imdb_clean)
```

# Regression Models

## Ordinary Linear Regression Model

### Creation of CV Folds

```{r}
data_cv10 <- vfold_cv(imdb_clean, v = 10)
```

### Model Spec, Recipes, and Workflows

```{r}
# Model Spec

lm_spec <-
    linear_reg() %>% 
    set_engine(engine = 'lm') %>% 
    set_mode('regression')

# Recipe

full_lm_rec <- recipe(Gross ~ Runtime + IMDB_Rating + Meta_score + 
                   No_of_Votes + Genre_1, data = imdb_clean) %>%
    step_nzv(all_predictors()) %>% 
    step_normalize(all_numeric_predictors()) %>% 
    step_dummy(all_nominal_predictors()) %>%
    step_naomit(Gross, Runtime, IMDB_Rating, Meta_score, No_of_Votes)

# Workflow

full_lm_wf <- workflow() %>%
    add_recipe(full_lm_rec) %>%
    add_model(lm_spec)
```

### Fit Full Model

```{r}
# All predictors included

full_lm_model <- fit(full_lm_wf, data = imdb_clean)

full_lm_model %>% tidy()
```

### Obtain Evaluation Metrics for Full Model

```{r}
# Cross Validation Performed (10 folds)

full_lm_modelcv <- fit_resamples(full_lm_wf, resamples = data_cv10, metrics = metric_set(rmse, rsq, mae))

full_lm_modelcv %>%
  collect_metrics()
```

## Perform LASSO for Subset Selection (Regression Model)

### Model Spec, Recipes, and Workflow

```{r}
# Lasso Model Spec with tune

lm_lasso_spec_tune <- 
  linear_reg() %>%
  set_args(mixture = 1, penalty = tune()) %>%   # mixture = 1 indicates Lasso
  set_engine(engine = 'glmnet') %>%             
  set_mode('regression')

# Recipe

data_rec_lasso <- recipe(Gross ~ Runtime + IMDB_Rating + Meta_score + 
                   No_of_Votes + Genre_1, data = imdb_clean) %>%
    step_nzv(all_predictors()) %>%                # removes variables with the same value (don't want duplicates)
    step_novel(all_nominal_predictors()) %>%      # important if you have rare categorical variables 
    step_normalize(all_numeric_predictors()) %>%  # standardization important step for LASSO
    step_dummy(all_nominal_predictors()) %>%      # creates indicator variables for categorical variables
    step_naomit(Gross, Runtime, IMDB_Rating,      # omit any NA values
                Meta_score, No_of_Votes)

# Workflow

lasso_wf_tune <- workflow() %>% 
  add_recipe(data_rec_lasso) %>%
  add_model(lm_lasso_spec_tune) 
```

### Tune Model and Cross Validation

```{r}
# Tune Model (to select best Lambda penalty) -- via Cross Validation

penalty_grid <- grid_regular(
  penalty(range = c(-3, 1)),
  levels = 30)

tune_res <- tune_grid(
  lasso_wf_tune, 
  resamples = data_cv10, 
  metrics = metric_set(rmse, mae),
  grid = penalty_grid 
)

# Visualize Model Evaluation Metrics from Tuning

autoplot(tune_res) + theme_classic()

# Collect CV Metrics and Select Best Model

# Summarize Model Evaluation Metrics (CV)
lasso_mod <- collect_metrics(tune_res) %>%
  filter(.metric == 'rmse') %>%
  select(penalty, rmse = mean)

# Choose penalty value
best_penalty <- select_best(tune_res, metric = 'rmse')

lasso_mod
```

### Fit Final LASSO Model

```{r}
# Fit Final Model -- Subset of Variables

final_wf <- finalize_workflow(lasso_wf_tune, best_penalty) # incorporates penalty value to workflow

final_fit <- fit(final_wf, data = imdb_clean)

tidy(final_fit)

# Final ("best") model predictors and coefficients

final_fit %>% tidy() %>% filter(estimate != 0)
```

### Obtain Evaluation Metrics for Lasso Model

```{r}
# Cross Validation Performed (10 folds)

lasso_fit <- fit_resamples(final_wf, resamples = data_cv10, metrics = metric_set(rmse, rsq, mae))

lasso_fit %>%
  collect_metrics()
```

### Visualize Residuals for LASSO Model

```{r}
lasso_mod_out <- final_fit %>%
    predict(new_data = imdb_clean) %>%
    bind_cols(imdb_clean) %>%
    mutate(resid = Gross - .pred)

ggplot(lasso_mod_out, aes(x = .pred, y = resid)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") + 
    theme_classic()

ggplot(lasso_mod_out, aes(x = Runtime, y = resid)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") + 
    theme_classic()

ggplot(lasso_mod_out, aes(x = IMDB_Rating, y = resid)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") + 
    theme_classic()

ggplot(lasso_mod_out, aes(x = No_of_Votes, y = resid)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") + 
    theme_classic()
```

## GAM with Splines (TidyModels)

### Build the GAM

```{r}
# Build the GAM

gam_spec <- 
  gen_additive_mod() %>%
  set_engine(engine = 'mgcv') %>%
  set_mode('regression')

gam_mod1 <- fit(gam_spec,
    Gross ~ s(Runtime, k = 20) + s(IMDB_Rating) + Meta_score + s(No_of_Votes) + Genre_1,
    data = imdb_clean 
)

```

### Run Diagnostics

````{r}
# Diagnostics: Check to see if the number of knots is large enough (if p-value is low, increase number of knots)

gam_mod1 %>% pluck('fit') %>% mgcv::gam.check()
```

```{r}
# Diagnostics: Check to see if the number of knots is large enough

gam_mod1 %>% pluck('fit') %>% summary() 
```

### Visualizations for Non-Linear Functions

```{r}
# Visualize: Look at the estimated non-linear functions

gam_mod1 %>% pluck('fit') %>% plot(main = "Non-Linear Visualizations")
```

### Obtain Evaluation Metrics for GAM1

```{r}
gam1_output <- gam_mod1%>% 
    predict(new_data = imdb_clean) %>%
    bind_cols(imdb_clean) %>%
    mutate(resid = Gross - .pred)

gam1_output %>%
    rmse(truth = Gross, estimate = .pred)

gam1_output %>%
    rsq(truth = Gross, estimate = .pred)

# Note: Unable to obtain std error for eval metrics bc can't perform cross validation w/ TidyModels
```

## GAM with Splines (Recipe)

### Build the GAM

```{r}
spline_rec <- recipe(Gross ~ Runtime + IMDB_Rating + Meta_score + 
                   No_of_Votes + Genre_1, data = imdb_clean) %>%
    step_nzv(all_predictors()) %>% 
    step_normalize(all_numeric_predictors()) %>% 
    step_dummy(all_nominal_predictors()) %>%
    step_naomit(Gross) %>%
    step_ns(Runtime, deg_free = 20) %>%
    step_ns(No_of_Votes, deg_free = 10) %>%
    step_ns(IMDB_Rating, deg_free = 10)
# Chose degrees of freedom to attempt to match # of knots from TidyModels
# May not be an exact match, but it's close

spline_rec %>% prep(imdb_clean) %>% juice()
```

```{r}
# Build the GAM

lm_spec_gam <-
  linear_reg() %>%
  set_engine(engine = 'lm') %>%
  set_mode('regression')

spline_wf <- workflow() %>%
    add_model(lm_spec) %>%
    add_recipe(spline_rec)

GAM2_fit <- fit(spline_wf, data = imdb_clean)

tidy(GAM2_fit)
```

### Obtain Evaluation Metrics for GAM2

```{r}
# Cross Validation Performed (10 folds)

cv_output_spline2 <- fit_resamples( 
  spline_wf, # workflow
  resamples = data_cv10, # cv folds
  metrics = metric_set(mae,rmse,rsq)
)

cv_output_spline2 %>% collect_metrics()
```

## Compare GAMs

```{r}
# Compare RMSE values --> pick model with lower RMSE

gam1_output %>%
    rmse(truth = Gross, estimate = .pred)

cv_output_spline2 %>% collect_metrics()
```

The GAM created using TidyModels performs better than the recipe GAM based on RMSE. Likely has to do with the degrees of freedom of the splines. However, it is worth noting that we were unable to perform cross validation on the TidyModels GAM.

### Visualize Residuals for Final GAM (GAM1)

```{r}
# Visualize Residuals for Final GAM (TidyModels GAM)

ggplot(gam1_output, aes(x = .pred, y = resid)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") + 
    labs(x = "Prediction",
         y = "Residuals", 
         title = "Residuals vs. Predictions") +
    theme_classic()

ggplot(gam1_output, aes(x = Runtime, y = resid)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red")  + 
    labs(x = "Runtime",
         y = "Residuals", 
         title = "Residuals vs. Runtime") +
    theme_classic()

ggplot(gam1_output, aes(x = IMDB_Rating, y = resid)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red")  + 
    labs(x = "IMDB Rating",
         y = "Residuals", 
         title = "Residuals vs. IMDB Rating") +
    theme_classic()

ggplot(gam1_output, aes(x = No_of_Votes, y = resid)) +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red")  + 
    labs(x = "Number of Votes",
         y = "Residuals", 
         title = "Residuals vs. Number of Votes") +
    theme_classic()
```

There does not appear to be any significant bias after analyzing the residuals. The trendline for Residuals vs. Runtime trends up at the end points since there are only a few cases with very short and very long runtimes.

## Compare All Regression Model Performance

```{r}
# Eval Metrics from CV

full_lm_modelcv %>% collect_metrics() # OLS
```

```{r}
# Eval Metrics from CV

lasso_fit %>% collect_metrics() # LASSO
```

```{r}
# Eval Metrics from CV

gam1_output %>%
    rmse(truth = Gross, estimate = .pred) # GAM
```

The GAM with Splines obtained using TidyModels performs the best with the lowest RMSE value (approx. 1.3 compared to approx. 1.4 for the others).

# Classification Models

## Create New Dataset for Classification

```{r}
# Remove log transformation from Gross

imdb_class <- imdb_top_1000 %>%
  cSplit("Genre", sep = ",", direction = "wide") %>%
  mutate(Gross = Revenue-Budget)

runtime_clean <- imdb_top_1000$Runtime %>%
  str_replace(" min", "") %>%
  as.numeric()

imdb_class$Runtime <- runtime_clean

imdb_class <- imdb_class %>%
  drop_na(Gross, Budget)

imdb_clean_class <- imdb_class %>%
  mutate(success_ratio = Revenue/Budget) %>%
  mutate(flop = as.factor(ifelse(success_ratio > 2, 'FALSE', 'TRUE'))) %>%
  drop_na(flop, No_of_Votes, Runtime, IMDB_Rating, Meta_score, Genre_1)
```

## Random Forests Model

### Model Spec, Recipe, and Workflow

```{r}
# Model Specification
rf_spec <- rand_forest() %>%
  set_engine(engine = 'ranger') %>% 
  set_args(trees = 1000, # Number of trees
           min_n = NULL,
           probability = FALSE,
           importance = 'impurity') %>%
  set_mode('classification')

# Recipe
data_rec <- recipe(flop ~ No_of_Votes + Runtime + IMDB_Rating + Meta_score + Genre_1,
                   data = imdb_clean_class) %>%
  step_naomit(flop, No_of_Votes, Runtime, IMDB_Rating, Meta_score, Genre_1)

# Create Workflows to Test Various Values for mtry

 data_wf_mtry3 <- workflow() %>%
  add_model(rf_spec %>% set_args(mtry = 3)) %>%
  add_recipe(data_rec)

data_wf_mtry4 <- workflow() %>%
  add_model(rf_spec %>% set_args(mtry = 4)) %>%
  add_recipe(data_rec)

data_wf_mtry5 <- workflow() %>%
  add_model(rf_spec %>% set_args(mtry = 5)) %>%
  add_recipe(data_rec)
```

### Fit Models with Various Values for mtry

```{r}
# Fit Models (for each mtry value)

set.seed(123)
data_fit_mtry3 <- fit(data_wf_mtry3, data = imdb_clean_class)

set.seed(123) 
data_fit_mtry4 <- fit(data_wf_mtry4, data = imdb_clean_class)

set.seed(123)
data_fit_mtry5 <- fit(data_wf_mtry5, data = imdb_clean_class)
```

### Obtain OOB Predictions and Evaluation Metrics to Choose mtry Value

```{r}
# Custom Function to get OOB predictions, true observed outcomes and add a user-provided model label

rf_OOB_output <- function(fit_model, model_label, truth){
    tibble(
          .pred_class = fit_model %>% extract_fit_engine() %>% pluck('predictions'), #OOB predictions
          flop = truth,
          model = model_label
      )
}
```

```{r}
# Evaluate OOB Metrics

data_rf_OOB_output <- bind_rows(
    rf_OOB_output(data_fit_mtry3,3, imdb_clean_class %>% pull(flop)),
    rf_OOB_output(data_fit_mtry4,4, imdb_clean_class %>% pull(flop)),
    rf_OOB_output(data_fit_mtry5,5, imdb_clean_class %>% pull(flop))
)

# OOB Accuracy for Each mtry Value

data_rf_OOB_output %>% 
    group_by(model) %>%
    accuracy(truth = flop, estimate = .pred_class)

data_rf_OOB_output %>% 
    group_by(model) %>%
    sensitivity(truth = flop, estimate = .pred_class)
# mtry = 3 and 4 tied for best sensitivity

# Accuracy vs. mtry Plot

data_rf_OOB_output %>% 
  group_by(model) %>%
  accuracy(truth = flop, estimate = .pred_class) %>%
  mutate(mtry = as.numeric(stringr::str_replace(model,'mtry',''))) %>%
  ggplot(aes(x = mtry, y = .estimate )) + 
  geom_point() +
  geom_line() +
  theme_classic()
# Select mtry = 4 based on accuracy
```

### Evaluating Selected Model - Confusion Matrix and Evaluation Metrics

```{r}
# Confusion Matrix

rf_OOB_output(data_fit_mtry4,4, imdb_clean_class %>% pull(flop)) %>%
    conf_mat(truth = flop, estimate= .pred_class)

# Sensitivity, Specificity, and Accuracy

rf_OOB_output(data_fit_mtry4,4, imdb_clean_class %>% pull(flop)) %>%
  sensitivity(truth = flop, estimate= .pred_class)

rf_OOB_output(data_fit_mtry4,4, imdb_clean_class %>% pull(flop)) %>%
  specificity(truth = flop, estimate= .pred_class)

rf_OOB_output(data_fit_mtry4,4, imdb_clean_class %>% pull(flop)) %>%
  accuracy(truth = flop, estimate= .pred_class)
```

### Variable Importance

#### Impurity

```{r}
# Impurity

model_output <-data_fit_mtry4 %>% 
    extract_fit_engine()

model_output %>% 
    vip(num_features = 10) + theme_classic() #based on impurity, 10 meaning the top 10

model_output %>% vip::vi() %>% head()
model_output %>% vip::vi() %>% tail()
```

#### Permuation

```{r}
# Permutation

model_output2 <- data_wf_mtry4 %>% 
  update_model(rf_spec %>% set_args(importance = "permutation")) %>% #based on permutation
  fit(data = imdb_clean_class) %>% 
    extract_fit_engine()

model_output2 %>% 
    vip(num_features = 10) + theme_classic()

model_output2 %>% vip::vi() %>% head()
model_output2 %>% vip::vi() %>% tail()
```

### Violin Graphs

```{r}
ggplot(imdb_clean_class, aes(x = flop, y = No_of_Votes)) +
    geom_violin() + theme_classic()

ggplot(imdb_clean_class, aes(x = flop, y = Runtime)) +
    geom_violin() + theme_classic()

ggplot(imdb_clean_class, aes(x = flop, y = IMDB_Rating)) +
    geom_violin() + theme_classic()

ggplot(imdb_clean_class, aes(x = flop, y = Meta_score)) +
    geom_violin() + theme_classic()
```

## Logistic Regression

### Model Spec, Recipe, and Workflow

```{r}
set.seed(123)

# Logistic Regression Model Spec
logistic_spec <- logistic_reg() %>%
    set_engine('glm') %>%
    set_mode('classification')

# Recipe
logistic_rec <- recipe(flop ~ No_of_Votes + Runtime + IMDB_Rating + Genre_1,
                   data = imdb_clean_class)

# Workflow (Recipe + Model) for Full Log Model
log_wf <- workflow() %>%
    add_recipe(logistic_rec) %>%
    add_model(logistic_spec)
```

### Fit Model

```{r}
# Fit Model

log_fit <- fit(log_wf, data = imdb_clean_class)

tidy(log_fit)
```

### Add Variable for Odds Ratio

```{r}
log_fit %>% tidy() %>%
  mutate(OR = exp(estimate))
```

### Cross Validation and Evaluation Metrics

```{r}
# Creation of CV Folds

data_cv10_class <- vfold_cv(imdb_clean_class, v = 10)
```

```{r}
# Cross Validation Performed (10 folds)

log_modelcv <- fit_resamples(log_wf, resamples = data_cv10_class, metrics = metric_set(accuracy,sens,yardstick::spec))

log_modelcv %>%
  collect_metrics()
```

### Picking Threshold

#### Boxplots

```{r}
final_output <- log_fit %>% predict(new_data = imdb_clean_class, type='prob') %>% bind_cols(imdb_clean_class)

final_output %>%
  ggplot(aes(x = flop, y = .pred_TRUE)) +
  geom_boxplot()
```

#### ROC Curve

```{r}
# Use soft predictions
final_output %>%
    roc_curve(flop,.pred_TRUE,event_level = 'second') %>%
    autoplot()
```

#### J Index vs. Threshold

```{r}
# Thresholds in terms of reference level

threshold_output <- final_output %>%
    threshold_perf(truth = flop, estimate = .pred_FALSE, thresholds = seq(0,1,by=.01))

# J-index v. Threshold for no flop

threshold_output %>%
    filter(.metric == 'j_index') %>%
    ggplot(aes(x = .threshold, y = .estimate)) +
    geom_line() +
    labs(y = 'J-index', x = 'threshold') +
    theme_classic()
```

```{r}
threshold_output %>%
    filter(.metric == 'j_index') %>%
    arrange(desc(.estimate))
```

#### Distance vs. Threshold

```{r}
# Distance vs. Threshold

threshold_output %>%
    filter(.metric == 'distance') %>%
    ggplot(aes(x = .threshold, y = .estimate)) +
    geom_line() +
    labs(y = 'Distance', x = 'threshold') +
    theme_classic()
```

```{r}
threshold_output %>%
    filter(.metric == 'distance') %>%
    arrange(.estimate)
```

### Obtain Evaluation Metrics for Logistic Regression Model with Thresholds

```{r}
# To determine final threshold

log_metrics <- metric_set(accuracy,sens,yardstick::spec)

# Compare Eval Metrics

final_output %>%
    mutate(.pred_class = make_two_class_pred(.pred_FALSE, levels(flop), threshold = .78)) %>%
    log_metrics(truth = flop, estimate = .pred_class, event_level = 'second')

final_output %>%
    mutate(.pred_class = make_two_class_pred(.pred_FALSE, levels(flop), threshold = .71)) %>%
    log_metrics(truth = flop, estimate = .pred_class, event_level = 'second')

# Compare Confusion Matrices

final_output %>%
  mutate(.pred_class = make_two_class_pred(.pred_FALSE, levels(flop), threshold = .78)) %>%
  conf_mat(truth = flop, estimate = .pred_class)

final_output %>%
  mutate(.pred_class = make_two_class_pred(.pred_FALSE, levels(flop), threshold = .71)) %>%
  conf_mat(truth = flop, estimate = .pred_class)
```

We chose the threshold of 0.78 for our final model because we are prioritizing sensitivity over accuracy and specificity.

### Predictions

```{r}
predict(log_fit, new_data = data.frame(No_of_Votes = 10000, Runtime = 112, IMDB_Rating = 9.8,
                                        Genre_1 = "Drama"), type = "prob"
)
```

We manually performed the hard predictions with threshold = .78. Since .pred_FALSE = 0.56 which is lower than our threshold of 0.78, we would predict this example movie would flop (flop = TRUE).

# Unsupervised Learning - Clustering

## K-Means Clustering

### Preliminary Visualizations

```{r}
ggplot(imdb_clean, aes(x = Budget, y = Runtime)) + 
  geom_point() + theme_classic()

imdb_clean %>%
  filter(Budget > 1) %>% 
  ggplot(aes(x = Budget, y = Gross)) + 
  geom_point() + 
  labs(x = "Budget in USD", 
       y = "Gross Profit in USD (Log Scale)", 
       title = "Preliminary Visualizations") +
  theme_classic()

ggplot(imdb_clean, aes(x = No_of_Votes, y = Runtime)) + 
  geom_point() + theme_classic()

ggplot(imdb_clean, aes(x = Gross, y = No_of_Votes)) + 
  geom_point() + theme_classic()
```

### Feature Selection

```{r}
imdb_sub <- imdb_clean %>%
    select(Budget, Gross)

set.seed(253)
```

### Determine Number of Clusters

```{r}
# Data-specific function to cluster and calculate total within-cluster SS
imdb_cluster_ss <- function(k){
    # Perform clustering
    kclust <- kmeans(scale(imdb_sub), centers = k)

    # Return the total within-cluster sum of squares
    return(kclust$tot.withinss)
}

tibble(
    k = 1:15,
    tot_wc_ss = purrr::map_dbl(1:15, imdb_cluster_ss)
) %>% 
    ggplot(aes(x = k, y = tot_wc_ss)) +
    geom_point() + 
    labs(x = "Number of clusters",y = 'Total within-cluster sum of squares') + 
    theme_classic()
```

### Select k = 8 Clusters

```{r}
kclust_k8 <- kmeans(scale(imdb_sub), centers = 8)

kclust_k8$cluster   # Display cluster assignments

imdb_clean <- imdb_clean %>%
    mutate(kclust_8 = factor(kclust_k8$cluster))
```

### Visualize Cluster Assignments

```{r}
# Visualize the cluster assignments on the original scatterplot
imdb_clean %>%
  ggplot(aes(x = Budget, y = Gross, color = kclust_8)) +
    geom_point() + theme_classic()
```

### Interpreting Clusters

#### Exploring Genre Breakdown

```{r}

# Count of Movies per Genre (Primary Genre)
imdb_clean %>%
  count(Genre_1)

# Count of Movies per Genre (Secondary Genre)
imdb_clean %>%
  count(Genre_2)

# Count of Movies per Genre (Overall Genre)
imdb_clean %>%
  count(New_Genre)
```

```{r}
# Genres vs Cluster

# How many of each Genre 1 in each cluster
imdb_clean %>%
  group_by(kclust_8) %>%
  count(Genre_1)

# How many of each Genre 2 in each cluster
imdb_clean %>%
  group_by(kclust_8) %>%
  count(Genre_2)

# How many movies in each cluster
imdb_clean%>%
  count(kclust_8)
```

#### Visualizations of Genres in Each Cluster

```{r}

# Genre 1
imdb_clean %>%
  ggplot(aes(x = kclust_8, fill = Genre_1)) +
    geom_bar(position = "fill") +
    labs(x = "Cluster") + 
    theme_classic()

# Genre 2
imdb_clean %>%
  ggplot(aes(x = kclust_8, fill = Genre_2)) +
    geom_bar(position = "fill") +
    labs(x = "Cluster") + 
    theme_classic()

# Overall Genre
imdb_clean %>%
  ggplot(aes(x = kclust_8, fill = New_Genre)) +
    geom_bar(position = "fill") +
    labs(x = "Cluster") + 
    theme_classic()

```

## Hierarchical Clustering

### Set up Clustering and Distance Matrix

```{r}
# Random subsample of 25 Movies
set.seed(253)

imdb_hc <- imdb_clean %>%
  slice_sample(n = 25) %>%
  filter(Budget != 0)

# Select the variables to be used in clustering
imdb_hc_sub <- imdb_hc %>%
  select(Gross, Budget)

imdb_hc_full <- imdb_clean %>%
  select(Gross, Budget) %>%
  filter(Budget > 1)

# Summary statistics for the variables
summary(imdb_hc_sub)

# Compute a distance matrix on the scaled data
dist_mat_scaled <- dist(scale(imdb_hc_sub))     # Subset Distance Matrix

dist_mat_full <- dist(scale(imdb_hc_full))      # Full Data Distance Matrix
```

### Perform Fusing Process

```{r}
imdb_hc_avg <- hclust(dist_mat_scaled, method = "average")    # Subset
imdb_full_avg <- hclust(dist_mat_full, method = "average")    # Full Data
```

### Visualize Dendrograms

```{r}
# Plot dendrogram on Subset
plot(imdb_hc_avg)
```

```{r}
# Adding Genre Labels

plot(imdb_hc_avg, labels = imdb_hc$Genre_1, main = "Movie Clusters", xlab = NULL)

plot(imdb_hc_avg, labels = paste(imdb_hc$Genre_1, imdb_hc$Genre_2))

plot(imdb_hc_avg, 
     main = "Visualizing Movie Clusters",
     labels = paste(imdb_hc$Genre_1, imdb_hc$Genre_2, imdb_hc$Genre_3), 
     hang = -1, 
     cex = 1)

plot(imdb_hc_avg, labels = paste(imdb_hc$New_Genre))
```

### Cutting the Tree (Choosing k)

```{r}
imdb_clean_clust <- imdb_clean %>%
  filter(Budget > 1) %>%
    mutate(
        hclust_num2 = factor(cutree(imdb_full_avg, k = 2)), # Cut into 2 clusters (k)
        hclust_num4 = factor(cutree(imdb_full_avg, k = 4)), # Cut into 4 clusters (k)
        hclust_num8 = factor(cutree(imdb_full_avg, k = 8)) # Cut into 8 clusters (k)
    )
```

```{r}
ggplot(imdb_clean_clust, aes(x = hclust_num2, fill = Genre_1)) +
  geom_bar(position = "fill") +
  labs(x = "Cluster", 
       y = "Proportion of Cluster", 
       title = "Selecting Number of Clusters (k)") + 
  theme_classic() + 
  theme(plot.title = element_text(size = 20, face = "bold"))

ggplot(imdb_clean_clust, aes(x = hclust_num4, fill = Genre_1)) +
  geom_bar(position = "fill") +
  labs(x = "Cluster", 
       y = "Proportion of Cluster", 
       title = "Selecting Number of Clusters (k)") + 
  theme_classic() + 
  theme(plot.title = element_text(size = 20, face = "bold"))

ggplot(imdb_clean_clust, aes(x = hclust_num8, fill = Genre_1)) +
  geom_bar(position = "fill") +
  labs(x = "Cluster", 
       y = "Proportion of Cluster", 
       title = "Selecting Number of Clusters (k)") + 
  theme_classic() + 
  theme(plot.title = element_text(size = 20, face = "bold"))
```

### Visualizing Genres in Final Clusters (Full Data)

```{r}
ggplot(imdb_clean_clust, aes(x = hclust_num2, fill = Genre_1)) +
  geom_bar(position = "fill") +
  labs(x = "Cluster", 
       y = "Proportion of Cluster", 
       title = "Selecting Number of Clusters (k)", 
       fill = "Genre 1") + 
  theme_classic() + 
  theme(plot.title = element_text(size = 20, face = "bold"))

ggplot(imdb_clean_clust, aes(x = hclust_num2, fill = New_Genre)) +
    geom_bar(position = "fill") +
    labs(x = "Cluster") + 
    theme_classic()

ggplot(imdb_clean_clust, aes(x = Budget, y = Gross, color = hclust_num2)) +
  geom_point() +
  labs(x = "Budget in USD", 
       y = "Gross Profit in USD (Log Scale)", 
       title = "Visualizing Clusters: Gross Profit vs. Budget",
       color = "Clusters") + 
  theme_classic() + 
  theme(plot.title = element_text(size = 20, face = "bold"))

ggplot(imdb_clean_clust, aes(x = hclust_num2, y = Budget)) +
  geom_boxplot() +
  labs(x = "Cluster", 
       y = "Budget in USD", 
       title = "Visualizing Clusters: Budget") + 
  theme_classic() + 
  theme(plot.title = element_text(size = 20, face = "bold"))

ggplot(imdb_clean_clust, aes(x = hclust_num2, y = Gross)) +
  geom_boxplot() +
  labs(x = "Cluster", 
       y = "Gross Profit in USD (Log Scale)", 
       title = "Visualizing Clusters: Gross Profit") + 
  theme_classic() + 
  theme(plot.title = element_text(size = 20, face = "bold"))

imdb_clean_clust %>%
  count(hclust_num2)

imdb_clean_clust %>%
  group_by(hclust_num2) %>%
  summarize(mean(Gross), sd(Gross), min(Gross), max(Gross),
            mean((Budget/100000)), sd((Budget/100000)), min((Budget/100000)), max((Budget/100000)))
```